---
title: "Poverty Analysis"
author: "AMDeines"
date: "1/18/2022"
output: html_document
---


# Analysis 
```{r load_mainData}
source("InFish_Auxillary.R")
#most recent output from "Poverty_CompileData.Rmd"
load(file="Output/MainData.RDATA")
str(MPI_FAO_Income)
```


Based on [working MS](https://tnc.box.com/s/8zlpzrojm6vl9dai4mn26st7t35pq6h1) as of 2022_01_19.

#Results
## Relationship between MPI and animal protein consumption   

 
###Contribution of freshwater fish in people’s diet in different levels of poverty (MPI)  

Animal protein consumed (g/person/day; y axis) ~ Multidimensional poverty (0-1, x axis) and size of points is the percentage freshwater fish makes up of the total animal protein consumed 

Data sources; Animal protein consumed is from ABC; MPI is from MPI Table 1 
```{r }
plot01 <-
  ggplot(
    data = MPI_FAO_Income %>%
      mutate(Protein_fwf = FAO_FreshwaterFishProtein_g.capita.day /
               FAO_TotalProtein_g.capita.day),
    aes(y = FAO_TotalProtein_g.capita.day, x = MPI, size = Protein_fwf)
  ) +
  expAnnotate(scpt="`Poverty_Analysis.Rmd`",
                figID="plot01")+
  geom_point()+
  ylab("Animal protein consumed (g/person/day)")+
  scale_size("Freshwater fish % of Protein")+
  expTheme

```
### Contribution of freshwater of freshwater fish in people’s diet in each region at different levels of poverty (MPI) and the trend/relationship between them.  

Animal protein consumed (g/person/day; y axis) ~ Multidimensional poverty (0-1, x axis) and size of points is the percentage freshwater fish makes up of the total and disaggregated by region (MPI table definition) and a trend line showing the regression between protein consumed and multidimensional poverty 

Data sources; Animal protein consumed is from ABC; MPI is from MPI Table 1 
```{r }
plot02 <-
  ggplot(
    data = MPI_FAO_Income %>%
      mutate(Protein_fwf = FAO_FreshwaterFishProtein_g.capita.day /
               FAO_TotalProtein_g.capita.day),
    aes(y = FAO_TotalProtein_g.capita.day, x = MPI, size = Protein_fwf)
  ) +
  expAnnotate(scpt="`Poverty_Analysis.Rmd`",
                figID="plot02")+
  geom_point()+
  geom_smooth(method="lm")+
  facet_wrap(~`World region`,scale="free")+
  ylab("Animal protein consumed (g/person/day)")+
  scale_size("Freshwater fish % of Protein")+
  expTheme
```
 

### Contribution of freshwater of freshwater fish in people’s diet in each region at different levels of NUTRITIONAL poverty (MPI) and the trend/relationship between them.  

Animal protein consumed (g/person/day; y axis) ~ Multidimensional poverty Nutrition (0-1, x axis) and size of points is the percentage freshwater fish makes up of the total and disaggregated by region (MPI table definition) and a trend line showing the regression between protein consumed and multidimensional poverty. 

Data sources; Animal protein consumed is from ABC; MPI is from MPI Table 1 
```{r }
# By POPULATION
plot03 <-
  ggplot(
    data = MPI_FAO_Income %>%
      mutate(Protein_fwf = FAO_FreshwaterFishProtein_g.capita.day /
               FAO_TotalProtein_g.capita.day),
    aes(y = FAO_TotalProtein_g.capita.day, 
        x = `MPI_Nutrition_%Population`, 
        size = Protein_fwf)
  ) +
  expAnnotate(scpt="`Poverty_Analysis.Rmd`",
                figID="plot03")+
  geom_point()+
  geom_smooth(method="lm")+
  facet_wrap(~`World region`,scale="free")+
  ylab("Animal protein consumed (g/person/day)")+
  xlab("MPI Nutrition Deprivation (% Population, Uncensored Headcounts)")+
  scale_size("Freshwater fish % of Protein")+
  expTheme

# By CONTRIBUTION
plot04 <-
  ggplot(
    data = MPI_FAO_Income %>%
      mutate(Protein_fwf = FAO_FreshwaterFishProtein_g.capita.day /
               FAO_TotalProtein_g.capita.day),
    aes(y = FAO_TotalProtein_g.capita.day, 
        x = `MPI_Nutrition_%Contribution`, 
        size = Protein_fwf)
  ) +
  expAnnotate(scpt="`Poverty_Analysis.Rmd`",
                figID="plot04")+
  geom_point()+
  geom_smooth(method="lm")+
  facet_wrap(~`World region`,scale="free")+
  ylab("Animal protein consumed (g/person/day)")+
  xlab("MPI Nutrition Deprivation (% Contribution)")+
  scale_size("Freshwater fish % of Protein")+
  expTheme
```
 

### Contribution of freshwater of freshwater fish in people’s diet in each region at different levels of conventional poverty measure and the trend/relationship between them.  

Animal protein consumed (g/person/day; y axis) ~ Percentage of population living on < US$1.90 a day (0-1, x axis) and size of points is the percentage freshwater fish makes up of the total and disaggregated by region (MPI table definition) and a trend line showing the regression between protein consumed and Poverty Gap 
```{r }
plot05 <-
  ggplot(
    data = MPI_FAO_Income %>%
      mutate(Protein_fwf = FAO_FreshwaterFishProtein_g.capita.day /
               FAO_TotalProtein_g.capita.day),
    aes(y = FAO_TotalProtein_g.capita.day, 
        x = WB_PovertyGap_1.90, 
        size = Protein_fwf)
  ) +
  expAnnotate(scpt="`Poverty_Analysis.Rmd`",
                figID="plot05")+
  geom_point()+
  geom_smooth(method="lm")+
  facet_wrap(~`World region`,scale="free")+
  ylab("Animal protein consumed (g/person/day)")+
  xlab("Poverty gap at $1.90 a day (%)")+
  scale_size("Freshwater fish % of Protein")+
  expTheme
```
 

### Contribution of freshwater of freshwater fish in people’s diet in Africa region at different levels of poverty (MPI) and the trend/relationship between them but controlled (in the loosest definition) for available of water to fish from. 

Only for Africa countries – show the above relationship but transformed to account for freshwater area / borders in a country.  
```{r }
plot06 <-"TBD"
```














# Old Stuff
## Nutrition Analysis
### Animal Protein~MPI
```{r AnimalProtein_by_MPI}
# Need to Join the FAO_protein data and MPI_raw data by Country and year
head(FAO_protein)
head(MPI_raw)
head(MPI_Deprivations)
head(MPI_headcounts)


# Unify Country Names using countrycode package
## Convert FAO to ISO as in the MPI data
## Check for matches, deal with manual missing matches
FAO_protein_1<-FAO_protein%>%
  filter(Year>=min(unlist(str_split(MPI_raw$`MPI data source_Year`,"-"))))%>% 
    ##filter by year to matches MPI data, and resolves some country dissolutions
  mutate(iso_country3=countrycode(Area.Code..FAO., origin = "fao", destination = "iso3c",
                      custom_match=  c(
                        "41"="CHN", # China, mainland
                        "151"="NA",  # Netherlands Antilles (former) dissolved by 2010 in FAO data
                        "206"="NA") # Sudan (former). South Sudan MPI data is 2010, Sudan is 2014          
  ))%>%
  group_by(iso_country3,Year)%>%
  summarise(Total_animal_protein=sum(Value,na.rm=TRUE),
            Total_FWF_protein=sum(Value[Item%in%"Freshwater Fish"],na.rm=TRUE),
            FWF_precent_Total=Total_FWF_protein/Total_animal_protein)
  ### Note, The Netherlands Antilles dissolved into several states, none of
  ### which appear to be in the MPI Data.  For Sudan (former), the 2010 FAO
  ### doesn't seem to have South Sudan.

# For each MPI country, grab the FAO data for the survey data year


MPI_raw_1<-MPI_raw%>%
  group_by( Country)%>%
  mutate(mpi_year_max= min(as.numeric(unlist(str_split(`MPI data source_Year`,"-")))))
  ## ("mpi_year_max") should be used because one would not expect a later
  ## poverty metric (say 2017) to have influenced an earlier fishing levels (say
  ## 2014), but rather fishing level in 2017 may have a signal of poverty
  ## metrics from previous years. "mpi_year_min" would be appropriate if the
  ## hypothesis was that fishing influenced poverty levels.


# Join the databases by Country and Year
Protein_MPI<-MPI_raw_1%>%
  left_join(MPI_Deprivations%>%
              select(`ISO country code`,
                     "Depriv_Health Nutrition"="Health Nutrition",
                     "Depriv_Living Standards Drinking water"="Living Standards Drinking water"))%>%
  left_join(MPI_headcounts%>%
              select(`ISO country code`,
                     "headcounts_Health Nutrition"="Health Nutrition",
                     "headcounts_Living Standards Drinking water"="Living Standards Drinking water"))%>%
  left_join(FAO_protein_1,
            by=c("ISO country code"="iso_country3","mpi_year_max"="Year") )


#plot
Protein_MPI_plot_01<-
  ggplot(data=Protein_MPI,
         aes(x=`Multidimensional poverty_Multidimensional Poverty Index`,
             y=`Total_animal_protein`))+
    expAnnotate(scpt=paste("'Poverty_FAOdata.Rmd'",sep=" "),
                figID="Protein_MPI_plot_01")+
  geom_point(aes(size=FWF_precent_Total*100,
                 color=`World region`))+
  #geom_text(aes(label=Country))+
  ylab("Total animal protein in diet (g/capita/day)")+
  xlab("Multidimensional Poverty Index (MPI)")+
  scale_size_continuous("FWF % Total")+
  #scale_y_log10()+
  #scale_x_log10()+
  expTheme
  

FWF_MPI_plot_02<-ggplot(data=Protein_MPI,
         aes(x=`Multidimensional poverty_Multidimensional Poverty Index`,
             y=FWF_precent_Total*100))+
    expAnnotate(scpt=paste("'Poverty_FAOdata.Rmd'",sep=" "),
                figID="FWF_MPI_plot_02")+
  geom_point(aes(color=`World region`))+
  #geom_text(aes(label=Country))+
  ylab("FWF % total protein in diet (g/capita/day)")+
  xlab("Multidimensional Poverty Index (MPI)")+
  #scale_y_log10()+
  #scale_x_log10()+
  expTheme+facet_wrap(~`World region`)+
  geom_text(aes(label=Country))


Compare with traditional income poverty metric. 

```


## uncategorized analyses

```{r plot}
#plot MPI~total production
FAO_10yr_prod<-FAO_MPI%>%filter(Fish_year%in%2009:2018)%>% #for ten-year average
 group_by(Country)%>%
 summarise(Prod_10yrAvg=mean(value_Total_t,na.rm = TRUE),
           frac_Capture_Avg=mean(frac_Capture,na.rm = TRUE),
           MPI=unique(MPI))

AMD01<-ggplot(data=FAO_MPI%>%filter(Fish_year==MPI_year),
       aes(x=value_Total_t,y=MPI,color=frac_Capture))+
  geom_point()+
  scale_y_log10()+
  scale_x_log10()+
  xlab("Total inland production in recent MPI survey year(capture+aquaculture t)")+
  scale_color_continuous("Fraction Capture")+
  ggtitle("DRAFT- Country level MPI by total inland fisheries production")+
  ggrepel::geom_text_repel(data=FAO_MPI%>%
                             filter(Fish_year==MPI_year,
                                    frac_Capture>.95),
            aes(label=Country),size=2)+
  theme(legend.position="bottom")
print(AMD01)
ggsave(plot=AMD01,filename="Output/Figures/AMD01_MPI by total inland fisheries.png",
       height=5,width=6,units = "in")


#plot MPI~consumption
FAO_10yr_cons<-FAO_MPI%>%
  group_by(Country)%>%
  summarise(Cons_10yrAvg=mean(value_Consumption,na.rm = TRUE),
            MPI=unique(MPI))

AMD02<-ggplot(data=FAO_MPI%>%filter(Fish_year==MPI_year),
       aes(x=value_Consumption,y=MPI))+
  geom_point()+
  scale_y_log10()+
  scale_x_log10()+
  xlab("Inland fisheries in recent MPI survey year")+
  ggtitle("DRAFT- Country level MPI by total inland fisheries consumption")
print(AMD02)

AMD03<-ggplot(data=FAO_MPI%>%filter(Fish_year==MPI_year), 
              aes(x=value_Total_t,y=value_Consumption,
                  color=frac_Capture,
                  size=MPI))+
  geom_point()+
  ylab("Consumption in recent MPI survey year")+
  xlab("Total production in recent MPI survey year")+
  scale_color_continuous("Fraction Capture")+
  scale_size_continuous("MPI")+
  scale_x_log10()+
  scale_y_log10()+
  ggtitle("DRAFT- Country consumption by total inland fisheries production")+
  ggrepel::geom_text_repel(data=FAO_MPI%>%
                             filter(Fish_year==MPI_year,
                                    value_Total_t<1e3),
            aes(label=Country),size=3)
print(AMD03)
  ##Production and consumption are rather linear, except at the lowest levels of
  ##production where consumption is abnormally high, though few data points.
ggsave(plot=AMD03,filename="Output/Figures/AMD03_inland fisheries production and consumption.png",
       height=5,width=7,units = "in")

```

An interesting aspect of AMD01, is that there appears to be a large separation in MPI based on the fraction of the production that is capture, where the higher fraction of capture appears to be correlated with higher MPI. 

```{r MPI_FracCap}

AMD04<-ggplot(data=FAO_10yr_prod,
       aes(color=Prod_10yrAvg,y=MPI,x=frac_Capture_Avg))+
  geom_point()+
  scale_y_log10()+
  scale_x_log10()+
  xlab("Fraction of total production that is Capture fishery")+
  scale_color_continuous("10 year average total production")+
  ggtitle("DRAFT- Country level MPI by total inland fisheries production")
print(AMD04)


#capture or Aquaculture by MPI
FAO_MPI_long<-FAO_MPI%>%
                filter(Fish_year==MPI_year)%>%
                pivot_longer(cols=c(value_Capture_t, value_Aquaculture_t),
                             names_to="Harvest mode",values_to="Harvest tons")%>%
                mutate(`Per_cap Harvest KG`=`Harvest tons`*1000/Population)

AMD05<-ggplot(data=FAO_MPI_long,
              aes(x=MPI,y=`Harvest tons`,color=`Harvest mode`))+
  geom_point()+
  scale_y_log10()+
  scale_x_log10()+
  xlab("MPI")+
  ylab("Production in recent MPI survey year")+
  geom_smooth(method="lm")+
  ggtitle("DRAFT- Country harvest by MPI and harvest mode")
print(AMD05)
ggsave(plot=AMD05,filename="Output/Figures/AMD05_Country harvest by MPI and harvest mode.png",
       height=5,width=6,units = "in")


FAO_lm_Capture<-lm(data=FAO_MPI_long,log(`Harvest tons`+1)~log(MPI),
                   subset=FAO_MPI_long$`Harvest mode`%in%"value_Capture_t")
summary(FAO_lm_Capture)
plot(FAO_lm_Capture)

FAO_lm_Aquaculture<-lm(data=FAO_MPI_long,log(`Harvest tons`+1)~log(MPI),
                   subset=!FAO_MPI_long$`Harvest mode`%in%"value_Capture_t")
summary(FAO_lm_Aquaculture)
plot(FAO_lm_Aquaculture)

FAO_lm_total<-lm(data=FAO_MPI_long,log(`Harvest tons`+1)~log(MPI)*`Harvest mode`,
                   subset=FAO_MPI_long$`Harvest tons`>0)
summary(FAO_lm_total)
par(mfrow=c(2,2))
plot(FAO_lm_total)

```

## Per Captia 
MPI at a per capita production
```{r per_capita}
#basically AMD05 production by MPI, but percaptia production
AMD06<-ggplot(data=FAO_MPI_long,
              aes(x=MPI,y=`Per_cap Harvest KG`,color=`Harvest mode`))+
  geom_point()+
  scale_y_log10()+
  scale_x_log10()+
  xlab("MPI")+
  ylab("KG per capita Production in recent MPI survey year")+
  geom_smooth(method="lm")+
  ggtitle("DRAFT- KG Per Capita Country harvest by MPI and harvest mode")
print(AMD06)
ggsave(plot=AMD06,filename="Output/Figures/AMD06_Country per-capita harvest by MPI and harvest mode.png",
       height=5,width=6,units = "in")

FAO_lm_Capture_percap<-lm(data=FAO_MPI_long,log(`Per_cap Harvest KG`)~log(MPI)*`Harvest mode`,
                          subset=FAO_MPI_long$`Per_cap Harvest KG`>0)
summary(FAO_lm_Capture_percap)
anova(FAO_lm_Capture_percap)
par(mfrow=c(2,2))
plot(FAO_lm_Capture_percap)
#Huge fan of the residuals.


```


## Nutrition 
Using "table 7" which isn't harmonized over time, but I've year matched when appropriate. 
```{r nutrition}
MPI_nutrition_0<- read_excel("Data/Table-7-All-MPI-Data-Since-2010-2020.xlsx",
                           range="A5:T409",col_names=FALSE)

#readable columns names bc the original source file is format bonkers. 
MPI_nutrition_names<-apply(MPI_nurtition_0[1:5,],2,function(i) paste(na.omit(i),collapse="_"))

#Clean up the nutrition data. 
MPI_nutrition<-MPI_nutrition_0%>%
  #add field namesand select relevant fields 
  slice(-c(1:5))%>%
  rename_at(names(MPI_nutrition_0), function(x)MPI_nutrition_names )%>%
  select(`Year MPI published`,
    year_of_survey=Year, Country, 
         Nutrition_percent=`Percentage of people who are poor and deprived in…_Health_Nutrition_% Population`)%>%
  #select the most recent survey years, which often span two years
  separate(year_of_survey,into=c("Year1","Year2"),fill="right")%>%
  pivot_longer(c(Year1, Year2), names_to = "name",values_to="Year_of_survey")%>%
  filter(!is.na(Nutrition_percent),!is.na(`Year_of_survey`))%>%
  select(-name)%>%
  mutate(Year_of_survey=as.numeric(Year_of_survey),
         Nutrition_percent =as.numeric(Nutrition_percent ),
         `Year MPI published`=as.numeric(`Year MPI published`))%>%
  group_by(Country)%>%
  filter(Year_of_survey==max(Year_of_survey),
         `Year MPI published`==max(`Year MPI published`))%>%
    ##assume use the most recent year of survey as reported in the most recent publication
  select(-`Year MPI published`)%>%
  #Unify country names to FAO  
  left_join(MPIlookup)%>%
  mutate(Land.Area= ifelse(is.na(Land.Area), Country, Land.Area))%>%
  #join FAO_MPI data to nutrition data. 
  right_join(FAO_MPI_long,by=c("Country","Year_of_survey"="MPI_year"))
  
AMD07<-ggplot(data=MPI_nutrition,
              aes(x=MPI,y=`Per_cap Harvest KG`,
                  color=`Harvest mode`,
                  size=Nutrition_percent))+
  geom_point()+
  scale_y_log10()+
  scale_x_log10()+
  xlab("MPI")+
  ylab("KG per capita Production in recent MPI survey year")+
  geom_smooth(method="lm")+
  ggtitle("DRAFT- KG Per Capita Country harvest by nutrition % and harvest mode ")
print(AMD07)


AMD08<-ggplot(data=MPI_nutrition,
              aes(x=Nutrition_percent,y=`Per_cap Harvest KG`,color=`Harvest mode`))+
  geom_point()+
  scale_y_log10()+
  scale_x_log10()+
  xlab("% population nutrition deprived")+
  ylab("KG per capita Production in recent MPI survey year")+
  geom_smooth(method="lm")+
  ggtitle("DRAFT- KG Per Capita Country harvest by nutrition % and harvest mode")
print(AMD08)
ggsave(plot=AMD08,filename="Output/Figures/AMD08_Country per-capita harvest by nutrition and harvest mode.png",
       height=5,width=6,units = "in")

FAO_lm_percap_nut<-lm(data=MPI_nutrition,log(`Per_cap Harvest KG`)~log(Nutrition_percent)*`Harvest mode`,
                          subset=MPI_nutrition$`Per_cap Harvest KG`>0)
summary(FAO_lm_percap_nut)
anova(FAO_lm_percap_nut)
par(mfrow=c(2,2))
plot(FAO_lm_percap_nut)
#Huge fan of the residuals.


ggplot(data=MPI_nutrition,
              aes(x=Nutrition_percent,y=MPI))+
  geom_point()+scale_x_log10()+scale_y_log10()

summary(lm(log(MPI)~log(Nutrition_percent),data=MPI_nutrition))

#Note that % population of nutrition deprived explains like 95% of MPI
#even though it's weight is only 1/6. 

```

plot diet data - "PercentProteinFWFish - FAOSTAT.xlsx" 