---
title: "Poverty Index and InFish"
author: "AMDeines"
date: "9/3/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

source("InFish_Auxillary.R")
library(readxl)
library(countrycode)
```

# Data 
Project data is stored & Shared via [TNC's box](https://tnc.box.com/s/6vs4fbq2paeysla5o5zbyrbi5dvjtfra). 
Connecting to Box [using boxr](https://r-box.github.io/boxr/articles/boxr-app-interactive.html)


The data compiled here (assumed all by country (ISO) and year
+ FAO Production
  [X] Capture 
  [X] Aquaculture data  
+ FAO Consumption??? Do we do this?
+ MPI   
  [X] overall MPI
  [X] region (MPI table definition) 
  ++ Nutrition:
    [NA] "Censored headcount"(% population)
    [X] "Uncensored Headcounts"(% population)
    [X] "Contribut'n of Deprivations"(% Contribution)?
+ FAO Nutrition - Animal protein consumed 
  [X] freshwater fish (fwf)
  [X] total
+ Income Poverty - WorldBank
  [] $1.90 Poverty
  
  

The goal is to make a large dataset aligned by country and year.
Note that the "standard" for data headers is "Source_descriptionA_descriptionB_units"
mt= metric tons, tonnes


## FAO Capture and Aquculture 
Used online queries at FAO for [Capture Production](http://www.fao.org/fishery/statistics/global-capture-production/query/en) and [Aquaculture Production](http://www.fao.org/fishery/statistics/global-aquaculture-production/query/en) to download country-level data for 2008-2018 Inland fisheries as .csv files.

```{r FAO_Fish_data}
# FAO Capture and Aquaculture Production
# Updated 20220208
FAO_Capture_Culture<-read.csv("Data/capture and aquaculture production_FishStatJ_1950-2019.csv")
 # "https://tnc.box.com/s/xmloswen8414xcj1xf2nf4yvmm6fgaim"
 # I don't fully understand the Province of this file. It's not purely FishStat output
 # Like, who added the country code?

FAO_production<-FAO_Capture_Culture%>%
  pivot_longer(X1950:X2019)%>%  # makes it long
  mutate(Year=as.numeric(gsub("X","",name)))%>%
  pivot_wider(names_from = Detailed.production.source,values_from = value)%>%
  select(Country,ISO3.code,Year,
         FAO_Aquaculture_production_mt=`Aquaculture production (freshwater)`,
         FAO_Capture_production_mt=`Capture production`)

```

## FAO Nutrition data

```{r Nutrition_data}
# Consumption Data...not used?
FAO_consumption<-read.csv("Data/consumption_FishStatJ_1961-2017.csv")

FAO_consumption%>%
  pivot_longer(X1961:X2017)%>%  # makes it long
  mutate(Year=as.numeric(gsub("X","",name)))%>%
  select(Country,Year,
         FAO_Consumption_mt=value)

# Protein Data
FAO_protein_2009<-read_csv("Data/Animal_Protein_FAOSTAT_data_2008-2009.csv")#load old data
FAO_protein_2019<-read_csv("Data/Animal_Protein_FAOSTAT_data_2010-2019.csv")#Load new data



FAO_protein <-
  rbind(FAO_protein_2009, #bind data
        FAO_protein_2019) %>%
  select(Area,`Area Code`,Year,Item,Value)%>%
  group_by(Area, `Area Code`,Year) %>%
  mutate(FAO_TotalProtein_g.capita.day = sum(Value, na.rm = TRUE)) %>%
  pivot_wider(names_from = Item, values_from = Value) %>%
  select(Country = Area,
         `Area Code`,
         Year,
         `FAO_TotalProtein_g.capita.day`:`Meat, Aquatic Mammals`)%>%
  ungroup()%>%
  mutate(ISO3.code=countrycode(`Area Code`,origin="fao",destination="iso3c"))
#i Some values were not matched unambiguously: 41, 151, 206
# 
unique(FAO_protein[is.na(FAO_protein$ISO3.code),c("Country","Area Code")])
# Country                       Area.Code
#1 China, mainland                      41
#2 Netherlands Antilles (former)       151
#3 Sudan (former)                      206

```
There are currently 3 known problems with the country code matching, which aren't readily resolved.


## MPI data 
```{r MPI_data}

# Tab "National MPI Results" ####

MPI_header<-c( #Obviously, manually specified, b.c the file is very messy
"ISO country numeric code",
"ISO country code",
"Country",
"World region",
"MPI data source_Survey",
"MPI data source_Year",
"Multidimensional poverty_Multidimensional Poverty Index",#(MPI = H*A)_Range 0 to 1
"Multidimensional poverty_Headcount ratio",#: Population in multidimensional poverty (H)_% Population
"Multidimensional poverty_Intensity of deprivation", # among the poor (A)_Average % of weighted deprivations
"Multidimensional poverty_Vulnerable to poverty", #_% Population
"Multidimensional poverty_In severe poverty",#_% Population
"Multidimensional poverty_Inequality among the poor",# (variance)
"Destitution_Headcount ratio",#: Population in multidimensional destitution (HD)_% Population",
"Destitution_Proportion of MPI poor who are destitute",#
"Total population_Year of the survey",#_Thousands",
"Total population_Population 2018", #_Thousands",
"Total population_Population 2019",#_Thousands",
"Number of MPI poor people_Year of the survey",#_Thousands",
"Number of MPI poor people_Population 2018",#_Thousands",
"Number of MPI poor people_Population 2019",#_Thousands",
"Indicators included in the MPI",#_Total number of indicators included  (out of ten)_",
"Indicators included in the MPI_Indicator (s) missing"
)

MPI_raw<-read_excel("Data/Table-1-National-Results-MPI-2021.xlsx",
                sheet="1.1 National MPI Results",
                range="A10:V118",col_names = MPI_header)
  #This skips the messy header formatting, which is provided manually above


# Tab "1.6 Uncensored Headcounts" ####
## For poverty indicators "% of population" in
### 1)  "Nutrition" 
### 2) "Drinking Water" 
MPI_header_1.6<-c(
"ISO country numeric code",
"ISO country code",
"Country",
"World region",
"MPI data source_Survey",
"MPI data source_Year",
"Multidimensional poverty_Multidimensional Poverty Index",#(MPI = H*A)_Range 0 to 1
#Percentage of people who are deprived in….
"Health Nutrition",	
"Health Child mortality",
"Education Years of schooling",	
"Education School attendance",
"Living Standards Cooking fuel", 
"Living Standards Sanitation",
"Living Standards Drinking water",
"Living Standards Electricity",
"Living Standards Housing",
"Living Standards Assets",
"Indicators included in the MPI",#_Total number of indicators included  (out of ten)_",
"Indicators included in the MPI_Indicator (s) missing")

MPI_headcounts<-read_excel("Data/Table-1-National-Results-MPI-2021.xlsx",
                sheet="1.6 Uncensored Headcounts",
                range="A10:S118",col_names = MPI_header_1.6)



# Tab "1.3 Contribut'n of Deprivations" ####
## For "Percentage contribution of deprivations of each indicator to overall poverty..." of 
### 1) "Nutrition" 
### 2) "Drinking water"
MPI_header_1.3<-c(
"ISO country numeric code",
"ISO country code",
"Country",
"World region",
"MPI data source_Survey",
"MPI data source_Year",
"Multidimensional poverty_Multidimensional Poverty Index",#(MPI = H*A)_Range 0 to 1
#Percentage contribution of deprivations ...of each dimension to overall poverty
"Percentage contribution Health",
"Percentage contribution Education",	
"Percentage contribution Living Standards",
#Percentage contribution of deprivations of each indicator to overall poverty…
"Health Nutrition",	
"Health Child mortality",
"Education Years of schooling",	
"Education School attendance",
"Living Standards Cooking fuel", 
"Living Standards Sanitation",
"Living Standards Drinking water",
"Living Standards Electricity",
"Living Standards Housing",
"Living Standards Assets",
"Indicators included in the MPI",#_Total number of indicators included  (out of ten)_",
"Indicators included in the MPI_Indicator (s) missing")

MPI_Deprivations<-read_excel("Data/Table-1-National-Results-MPI-2021.xlsx",
                sheet="1.3 Contribut'n of Deprivations",
                range="A10:V118",col_names = MPI_header_1.3)


# Tab "1.4 MPI Results & Compl.[Compiled] Data" ####
MPI_header_1.4<-c(
"ISO country numeric code",
"ISO country code",
"Country",
"World region",
"MPI data source_Survey",
"MPI data source_Year",
"Multidimensional poverty_Multidimensional Poverty Index",#(MPI = H*A)_Range 0 to 1
"Population in multidimensional poverty", #H
"Intensity of deprivation among the poor",# (A)
"Income Poverty_1.90_Value",
"Income Poverty_1.90_Year",
"Income Poverty_3.10_Value",
"Income Poverty_3.10_Year",
"Income Poverty_NatPovertyLine_Value",
"Income Poverty_NatPovertyLine_Year",
"GNI per capita",
"Gini Index_Value",
"Gini Index_Year",
"HDI_Value",
"HDI_Category",
"Income category"
)

MPI_PovertyGap<-read_excel("Data/Table-1-National-Results-MPI-2021.xlsx",
                sheet="1.4 MPI Results & Compl. Data",
                range="A10:U118",col_names = MPI_header_1.4)

#Compile MPI data####
# from selected sheets We are only after the Nutrition component (for now)
MPI_raw # Turns out this isn't really needed, since the MPI is also in the other tables
MPI_headcounts
MPI_Deprivations
MPI_PovertyGap

MPI <- full_join(
    MPI_headcounts %>%
      select(Country, ISO3.code = `ISO country code`,`World region`, 
             `MPI data source_Year`,
             MPI="Multidimensional poverty_Multidimensional Poverty Index",
             MPI_Nutrition=`Health Nutrition`),
    MPI_Deprivations%>%
      select(Country, ISO3.code = `ISO country code`, 
             `MPI data source_Year`,
             MPI_Nutrition=`Health Nutrition`),
    by=c("Country","ISO3.code","MPI data source_Year"),
    suffix = c("_%Population", "_%Contribution"))%>%
  full_join(
    MPI_PovertyGap%>%
           select(Country, ISO3.code=`ISO country code`, 
           `MPI data source_Year`,
           `Income Poverty_1.90_Value`,`Income Poverty_1.90_Year`),
    by=c("Country","ISO3.code","MPI data source_Year"))%>%
  separate(`MPI data source_Year`,into=c(NA,"Latest_Year"),
           remove=FALSE, fill="left",convert=TRUE)%>%
  separate(`MPI data source_Year`,into=c("Earliest_Year",NA),
           remove=FALSE, fill="right",convert=TRUE)
  
MPI[MPI$Latest_Year-MPI$Earliest_Year>1,] #!!!CAUTION PNG has three years of data

#Check that the dimensions of the join make sense, should all be equal.
nrow(MPI_headcounts) #109
nrow(MPI_Deprivations)#109
nrow(MPI)#109
##Note also that the joining of the dates confirms these are just
##repeated in the tables

qplot(data=MPI,y=`MPI_Nutrition_%Population`,x=`MPI_Nutrition_%Contribution`)
##The nutrition "_%Population", "_%Contribution" are pretty unrealated. 

```

## Compile Data 
```{r compile}
# The MPI data set is the one we are trying to match to, so, join to that one.
# Do a join for each of the possible years in the MPI, latest first...then the rest

# FAO PRODUCTION DATA ####
# Do a join for each of the possible years in the MPI, latest first...then the rest
#All the matches to the lastest Year
MPI_FAO_Production_Latest <- inner_join(MPI,
                                        FAO_production%>%mutate(FAO_production_year=Year),
                                        by = c("ISO3.code", "Latest_Year" = "Year"))

#All the matches to the Earlist Year Not in Latest year
MPI_FAO_Production_Earlist <- left_join(
  MPI[!MPI$ISO3.code%in%MPI_FAO_Production_Latest$ISO3.code,],
  FAO_production%>%mutate(FAO_production_year=Year),
  by = c("ISO3.code", "Earliest_Year" = "Year"))

#Combine the latest and Earliest
MPI_FAO_Production<-bind_rows(MPI_FAO_Production_Latest,MPI_FAO_Production_Earlist)%>%
  select(-c(Country.y),Country=Country.x)

#Check that any 3+year data sets (PNG) are included
MPI_FAO_Production [MPI_FAO_Production$ISO3.code %in% 
                      MPI$ISO3.code[MPI$Latest_Year-MPI$Earliest_Year>1],]
##All Set.


# FAO Nutrition DATA ####
# There's no point in dealing with "old" data not in MPI ~ 2009
##Not all the

#All the matches to the lastest Year
MPI_FAO_protein_Latest <- inner_join(MPI_FAO_Production,
                                        FAO_protein%>%mutate(FAO_protein_year=Year),
                                        by = c("ISO3.code", "Latest_Year" = "Year"))

#All the matches to the Earlist Year Not in Latest year
MPI_FAO_protein_Earlist <- left_join(
  MPI_FAO_Production[!MPI_FAO_Production$ISO3.code%in%MPI_FAO_protein_Latest$ISO3.code,],
  FAO_protein%>%mutate(FAO_protein_year=Year),
  by = c("ISO3.code", "Earliest_Year" = "Year"))

#Combine the latest and Earliest
MPI_FAO<-bind_rows(MPI_FAO_protein_Latest,MPI_FAO_protein_Earlist)%>%
  select(-c(Country.y,`Area Code`),Country=Country.x)

#Check that any 3+year data sets (PNG) are included
as.data.frame(MPI_FAO[MPI_FAO$ISO3.code %in% 
                      MPI$ISO3.code[MPI$Latest_Year-MPI$Earliest_Year>1],])
# All Set 


```



```{r write}

#Write out compiled data
write.csv(MPI_FAO,file=paste0("Output/MainData_",Sys.Date(),".csv"))

# This values get's added to the Box account, 
# Then Sui is manually adding protien data

# Save for other analysis
save(MPI_FAO,file="Output/MainData.RDATA")
```