---
title: "Poverty Analysis"
author: "AMDeines"
date: "1/18/2022"
output: html_document
---


# Analysis 
```{r load_mainData}
source("InFish_Auxillary.R")
#most recent output from "Poverty_CompileData.Rmd"
load(file="Output/MainData.RDATA")
str(MPI_FAO_Income)
```


Based on [working MS](https://tnc.box.com/s/8zlpzrojm6vl9dai4mn26st7t35pq6h1) as of 2022_01_19.

#Results
## Relationship between MPI and animal protein consumption   

 
###Contribution of freshwater fish in people’s diet in different levels of poverty (MPI)  

Animal protein consumed (g/person/day; y axis) ~ Multidimensional poverty (0-1, x axis) and size of points is the percentage freshwater fish makes up of the total animal protein consumed 

Data sources; Animal protein consumed is from ABC; MPI is from MPI Table 1 
```{r }
plot01 <-
  ggplot(
    data = MPI_FAO_Income %>%
      mutate(Protein_fwf = FAO_FreshwaterFishProtein_g.capita.day /
               FAO_TotalProtein_g.capita.day),
    aes(y = FAO_TotalProtein_g.capita.day, x = MPI, size = Protein_fwf)
  ) +
  expAnnotate(scpt="`Poverty_Analysis.Rmd`",
                figID="plot01")+
  geom_point()+
  ylab("Animal protein consumed (g/person/day)")+
  scale_size("Freshwater fish % of Protein")+
  expTheme

```


### Contribution of freshwater of freshwater fish in people’s diet in each region at different levels of poverty (MPI) and the trend/relationship between them.  

Freshwater fish as a % of animal protein (%; y axis) ~ Multidimensional poverty (0-1, x axis) and size of points is the percentage freshwater fish makes up of the total and disaggregated by region (MPI table definition) and a trend line showing the regression between fish as a % of animal proteinand multidimensional poverty


Data sources; Animal protein consumed is from ABC; MPI is from MPI Table 1 
```{r }
plot02 <-
  ggplot(
    data = MPI_FAO_Income %>%
      mutate(Protein_fwf = FAO_FreshwaterFishProtein_g.capita.day /
               FAO_TotalProtein_g.capita.day),
    aes(y =Protein_fwf , x = MPI, size = FAO_TotalProtein_g.capita.day) #
  ) +
  expAnnotate(scpt="`Poverty_Analysis.Rmd`",
                figID="plot02")+
  geom_point()+
  geom_smooth(method="lm", show.legend = FALSE)+
  facet_wrap(~`World region`,scale="free")+
  ylab("Freshwater fish % of Protein")+ #
  scale_size("Animal protein consumed (g/person/day)")+
  expTheme


plot07<-
  ggplot(
    data = MPI_FAO_Income %>%
      mutate(Protein_fwf = FAO_FreshwaterFishProtein_g.capita.day /
               FAO_TotalProtein_g.capita.day)%>%
      filter(`World region`%in%"Sub-Saharan Africa"),
    aes(y =Protein_fwf , x = MPI,color=FAO_TotalProtein_g.capita.day) #
  ) +
  expAnnotate(scpt="`Poverty_Analysis.Rmd`",
                figID="plot07")+
  geom_point()+
  ggrepel::geom_text_repel(aes(label=Country))+
  ylab("Freshwater fish % of Protein")+ #
  expTheme




```
 

### Contribution of freshwater of freshwater fish in people’s diet in each region at different levels of NUTRITIONAL poverty (MPI) and the trend/relationship between them.  

Animal protein consumed (g/person/day; y axis) ~ Multidimensional poverty Nutrition (0-1, x axis) and size of points is the percentage freshwater fish makes up of the total and disaggregated by region (MPI table definition) and a trend line showing the regression between protein consumed and multidimensional poverty. 

Data sources; Animal protein consumed is from ABC; MPI is from MPI Table 1 
```{r }
# By POPULATION
plot03 <-
  ggplot(
    data = MPI_FAO_Income %>%
      mutate(Protein_fwf = FAO_FreshwaterFishProtein_g.capita.day /
               FAO_TotalProtein_g.capita.day),
    aes(size = FAO_TotalProtein_g.capita.day, 
        x = `MPI_Nutrition_%Population`, 
        y = Protein_fwf)
  ) +
  expAnnotate(scpt="`Poverty_Analysis.Rmd`",
                figID="plot03")+
  geom_point()+
  geom_smooth(method="lm", show.legend = FALSE)+
  facet_wrap(~`World region`,scale="free")+
  ylab("Freshwater fish % of Protein")+
  xlab("MPI Nutrition Deprivation (% Population, Uncensored Headcounts)")+
  scale_size("Animal protein consumed (g/person/day)")+
  expTheme

# By CONTRIBUTION
plot04 <-
  ggplot(
    data = MPI_FAO_Income %>%
      mutate(Protein_fwf = FAO_FreshwaterFishProtein_g.capita.day /
               FAO_TotalProtein_g.capita.day),
    aes(size = FAO_TotalProtein_g.capita.day, 
        x = `MPI_Nutrition_%Contribution`, 
        y = Protein_fwf)
  ) +
  expAnnotate(scpt="`Poverty_Analysis.Rmd`",
                figID="plot04")+
  geom_point()+
  geom_smooth(method="lm", show.legend = FALSE)+
  facet_wrap(~`World region`,scale="free")+
  ylab("Freshwater fish % of Protein")+
  xlab("MPI Nutrition Deprivation (% Contribution)")+
  scale_size("Animal protein consumed (g/person/day)")+
  expTheme
```
 

### Contribution of freshwater of freshwater fish in people’s diet in each region at different levels of conventional poverty measure and the trend/relationship between them.  

Animal protein consumed (g/person/day; y axis) ~ Percentage of population living on < US$1.90 a day (0-1, x axis) and size of points is the percentage freshwater fish makes up of the total and disaggregated by region (MPI table definition) and a trend line showing the regression between protein consumed and Poverty Gap 
```{r }
plot05 <-
  ggplot(
    data = MPI_FAO_Income %>%
      mutate(Protein_fwf = FAO_FreshwaterFishProtein_g.capita.day /
               FAO_TotalProtein_g.capita.day),
    aes(size = FAO_TotalProtein_g.capita.day, 
        x = WB_PovertyGap_1.90, 
        y = Protein_fwf)
  ) +
  expAnnotate(scpt="`Poverty_Analysis.Rmd`",
                figID="plot05")+
  geom_point()+
  geom_smooth(method="lm", show.legend = FALSE)+
  facet_wrap(~`World region`,scale="free")+
  ylab("Freshwater fish % of Protein")+
  xlab("Poverty gap at $1.90 a day (%)")+
  scale_size("Animal protein consumed (g/person/day)")+
  expTheme
```
 

### Contribution of freshwater of freshwater fish in people’s diet in Africa region at different levels of poverty (MPI) and the trend/relationship between them but controlled (in the loosest definition) for available of water to fish from. 

Only for Africa countries – show the above relationship but transformed to account for freshwater area / borders in a country.  
```{r }
plot06 <-"TBD"
```












