---
title: "Poverty Index and InFish"
author: "AMDeines"
date: "9/3/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(lubridate)
library(readxl)
```

# Data 

## FAO Capture and Aquculture 
Used online queires at FAO for [Capture Production](http://www.fao.org/fishery/statistics/global-capture-production/query/en) and [Aquaculture Production](http://www.fao.org/fishery/statistics/global-aquaculture-production/query/en) to download country-level data for 2008-2018 Inland fisheries as .csv files.

```{r FAO_import}
FAO_capture<-read.csv("Data/CAPTURE_figis_20200903.csv")
FAO_aqua<-read.csv("Data/AQUA_figis_20200903.csv")

FAO_consumption<-read.csv("Data/FoodBalance_FWprodcons_2003-2013_MPI_del - FoodBalance_FWprodcons_2003-2013_MPI_del.csv")

```

```{r FAO_tidy}
FAO_capture_long<-FAO_capture%>%
  filter(Unit%in%"t")%>% #only use tonnes
  select(!starts_with("S_"),-Ocean.Area)%>% # ignores data flag for estimated data
  pivot_longer(X2008:X2018)%>%  # makes it long
  mutate(Year=as.numeric(gsub("X","",name)))%>%
  select(-name)#drop name field


FAO_aqua_long<-FAO_aqua%>%
  select(!starts_with("S_"),-Ocean.Area)%>% # ignores data flag for estimated data
  pivot_longer(X2008:X2018)%>%  # makes it long
  mutate(Year=as.numeric(gsub("X","",name)))%>%
  select(-name)#drop name field

FAO_consump_long<-FAO_consumption%>%
  select(!starts_with("prod_"),- Group.of.species..Name.,- Unit..Name.,-MPI.country.name)%>%
  pivot_longer(cons_2004:cons_2013,names_to="Year",values_to="value_Consumption")%>%
  mutate(Year=as.numeric(gsub("cons_","",Year)))
```


## MPI data 
```{r tidy_MPI}
MPI_raw<-read_excel("Data/mpi_table_2019_final_0.xlsx",
                sheet="Table 1",
                range="A5:E106")

MPI<-MPI_raw%>%select("Country",MPI="Value",MPI_year=`2007-2018`)%>%
  separate(MPI_year,sep=" ",into=c("mpi_years","mpi_flag"))%>% #separates years from flag
  separate(mpi_years,into=c("mpi_year_min","mpi_year_max"),fill="left",convert = TRUE)
    ## separates two years if given, with single years going into max allowing
    ## assumption to join to fish data by latest year of survey, consistent with
    ## the assumption that drivers of poverty metrics also are drivers of
    ## fishing levels, where the later of two poverty metric years
    ## ("mpi_year_max") should be used because one would not expect a later
    ## poverty metric (say 2017) to have influenced an earlier fishing levels
    ## (say 2014), but rather fishing level in 2017 may have a signal of poverty
    ## metrics from previous years. fill="right" and "mpi_year_min" would be
    ## appropriate if the hypothesis was that fishing influenced poverty levels.

```

## Population
http://www.fao.org/faostat/en/#data/OA 
```{r population_data}
#Add Population data here


```

## Compile Data 
```{r compile_FAOandMPI}
#Compare Country names across FAO capture, aquaculture, & Consumption data

#Capture & Aquaculture
table(FAO_capture_long$Land.Area%in%FAO_aqua_long$Land.Area)
table(FAO_aqua_long$Land.Area%in%FAO_capture_long$Land.Area)
## All Aquaculture Countries are in Capture data

#Capture & Consumption 
Capture_notin_Consumption<-unique(FAO_capture_long$Land.Area[
  !FAO_capture_long$Land.Area%in%FAO_consump_long$FAO.country.name])
Consumption_notin_Capture<-unique(FAO_consump_long$FAO.country.name[
  !FAO_consump_long$FAO.country.name%in%FAO_capture_long$Land.Area])

#rename Palistine in consumption data to match FAO data.
FAO_consump_long$FAO.country.name[FAO_consump_long$FAO.country.name%in%"Palestine, Occupied Tr."]<-
  "Palestine"

#Join all the FAO fish data
FAO<-full_join(FAO_capture_long,FAO_aqua_long,
               by=c("Land.Area","Year"),
               suffix = c("_Capture", "_Aquaculture"))%>% #join datasets
      rowwise()%>%
      mutate(value_Total=sum(value_Capture,value_Aquaculture,na.rm=TRUE),
             frac_Capture=value_Capture/value_Total)%>%
    full_join(FAO_consump_long,by=c("Land.Area"="FAO.country.name","Year"))

#Compare FAO Countries to MPI Countries
#MPI countries NOT in FAO
MPI_notin_FAO<-unique(MPI$Country[!MPI$Country%in%FAO$Land.Area])

#FAO countries NOT in MPI
FAO_notin_MPI<-unique(FAO$Land.Area[!FAO$Land.Area%in%MPI$Country])

#Manually create a lookup for MPI not in FAO
MPIlookup<-data.frame(Country=MPI_notin_FAO,
                      Land.Area=c("Bolivia (Plurinat.State)",
                                 "Congo, Dem. Rep. of the",
                                 "Côte d'Ivoire",
                                 "Eswatini",
                                 "Lao People's Dem. Rep.",
                                 "Moldova, Republic of",
                                 "Palestine",
                                 "Tanzania, United Rep. of"))
# FAO-MPI join using the lookup table
FAO_MPI<-MPI%>%
  left_join(MPIlookup)%>%
  mutate(Land.Area = ifelse(is.na(Land.Area), Country, Land.Area))%>%
  full_join(FAO,by="Land.Area")%>%
  select(Country=Land.Area,Fish_year=Year,MPI,MPI_year=mpi_year_max,value_Capture_t=value_Capture,
         value_Aquaculture_t=value_Aquaculture,value_Total_t=value_Total,
         frac_Capture, value_Consumption)
```


# Analysis 
```{r plot}
#plot MPI~total production
FAO_10yr_prod<-FAO_MPI%>%filter(Fish_year%in%2009:2018)%>% #for ten-year average
 group_by(Country)%>%
 summarise(Prod_10yrAvg=mean(value_Total_t,na.rm = TRUE),
           frac_Capture_Avg=mean(frac_Capture,na.rm = TRUE),
           MPI=unique(MPI))

AMD01<-ggplot(data=FAO_MPI%>%filter(Fish_year==MPI_year),
       aes(x=value_Total_t,y=MPI,color=frac_Capture))+
  geom_point()+
  scale_y_log10()+
  scale_x_log10()+
  xlab("Total inland production in recent MPI survey year(capture+aquaculture t)")+
  scale_color_continuous("Fraction Capture")+
  ggtitle("DRAFT- Country level MPI by total inland fisheries production")+
  ggrepel::geom_text_repel(data=FAO_MPI%>%
                             filter(Fish_year==MPI_year,
                                    frac_Capture>.95),
            aes(label=Country),size=2)+
  theme(legend.position="bottom")
ggsave(plot=AMD01,filename="Output/Figures/AMD01_MPI by total inland fisheries.png",
       height=5,width=6,units = "in")
print(AMD01)

#plot MPI~consumption
FAO_10yr_cons<-FAO_MPI%>%
  group_by(Country)%>%
  summarise(Cons_10yrAvg=mean(value_Consumption,na.rm = TRUE),
            MPI=unique(MPI))

AMD02<-ggplot(data=FAO_MPI%>%filter(Fish_year==MPI_year),
       aes(x=value_Consumption,y=MPI))+
  geom_point()+
  scale_y_log10()+
  scale_x_log10()+
  xlab("Inland fisheries in recent MPI survey year")+
  ggtitle("DRAFT- Country level MPI by total inland fisheries consumption")
print(AMD02)

AMD03<-ggplot(data=FAO_MPI%>%filter(Fish_year==MPI_year), 
              aes(x=value_Total_t,y=value_Consumption,
                  color=frac_Capture,
                  size=MPI))+
  geom_point()+
  ylab("Consumption in recent MPI survey year")+
  xlab("Total production in recent MPI survey year")+
  scale_color_continuous("Fraction Capture")+
  scale_size_continuous("MPI")+
  scale_x_log10()+
  scale_y_log10()+
  ggtitle("DRAFT- Country consumption by total inland fisheries production")+
  ggrepel::geom_text_repel(data=FAO_MPI%>%
                             filter(Fish_year==MPI_year,
                                    value_Total_t<1e3),
            aes(label=Country),size=3)
print(AMD03)
  ##Production and consumption are rather linear, except at the lowest levels of
  ##production where consumption is abnormally high, though few data points.
ggsave(plot=AMD03,filename="Output/Figures/AMD03_inland fisheries production and consumption.png",
       height=5,width=7,units = "in")

```

An interesting aspect of AMD01, is that there appears to be a large separation in MPI based on the fraction of the production that is capture, where the higher fraction of capture appears to be correlated with higher MPI. 

```{r MPI_FracCap}

AMD04<-ggplot(data=FAO_10yr_prod,
       aes(color=Prod_10yrAvg,y=MPI,x=frac_Capture_Avg))+
  geom_point()+
  scale_y_log10()+
  scale_x_log10()+
  xlab("Fraction of total production that is Capture fishery")+
  scale_color_continuous("10 year average total production")+
  ggtitle("DRAFT- Country level MPI by total inland fisheries production")
print(AMD04)


#capture or Aquaculture by MPI
FAO_MPI_long<-FAO_MPI%>%
                filter(Fish_year==MPI_year)%>%
                pivot_longer(cols=c(value_Capture_t, value_Aquaculture_t),
                             names_to="Harvest mode",values_to="Harvest tons")

AMD05<-ggplot(data=FAO_MPI_long,
              aes(x=MPI,y=`Harvest tons`,color=`Harvest mode`))+
  geom_point()+
  scale_y_log10()+
  scale_x_log10()+
  xlab("MPI")+
  ylab("Production in recent MPI survey year")+
  geom_smooth(method="lm")+
  ggtitle("DRAFT- Country harvest by MPI and harvest mode")
print(AMD05)
ggsave(plot=AMD05,filename="Output/Figures/AMD05_Country harvest by MPI and harvest mode.png",
       height=5,width=6,units = "in")


FAO_lm_Capture<-lm(data=FAO_MPI_long,log(`Harvest tons`+1)~log(MPI),
                   subset=FAO_MPI_long$`Harvest mode`%in%"value_Capture_t")
summary(FAO_lm_Capture)
plot(FAO_lm_Capture)

FAO_lm_Aquaculture<-lm(data=FAO_MPI_long,log(`Harvest tons`+1)~log(MPI),
                   subset=!FAO_MPI_long$`Harvest mode`%in%"value_Capture_t")
summary(FAO_lm_Aquaculture)
plot(FAO_lm_Aquaculture)



```